Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-24 ind_files/03_BD_processes.tex
Minor rephrasing

As in the one-dimensional case, we can go a little further if the noise is sufficiently weak. Let the deterministic trajectory obtained by solving \eqref{nD_det_limit} be given by $\boldsymbol{\alpha}(t)$.  We can once again track 

@@

&= \sum\limits_{k=1}^{m} y_k\left(a_i\frac{\partial w_i}{\partial x_k}\bigg{|}_{\mathbf{x}=\boldsymbol{\alpha}(t)}\right) + y_iw_i(\boldsymbol{\alpha}) + \mu\sum\limits_{k=1}^{m} y_k\left(\frac{\partial Q_i}{\partial x_k}(\mathbf{x})\bigg{|}_{\mathbf{x}=\boldsymbol{\alpha}(t)}\right)\\
&= y_iw_i(\boldsymbol{\alpha}) + a_iD_{\mathbf{y}}(w_i(\boldsymbol{\alpha})) + \mu D_{\mathbf{y}}(Q_i(\boldsymbol{\alpha}))\label{nD_WNA_directional_derivative_for_replicator_eqns}

@@

\frac{d}{dt}\mathbb{E}[y_i] = \underbrace{w_i(\boldsymbol{\alpha})\mathbb{E}[y_i]}_{\substack{\text{Current fitness of type $i$} \\ \text{at deterministic trajectory $\boldsymbol{\alpha}$} \\ \text{(scaled by expected density $\mathbb{E}[y_i]$)}}} + \underbrace{a_i\mathbb{E}[D_{\mathbf{y}}(w_i(\boldsymbol{\alpha}))]}_{\substack{\text{Expected change in fitness} \\ \text{ of type $i$ in going from $\boldsymbol{\alpha}$ to $\mathbf{y}$} \\ \text{(scaled by deterministic density $a_i$)}}} + \underbrace{\mu\mathbb{E}[D_{\mathbf{y}}(Q_i(\boldsymbol{\alpha}))]}_{\substack{\text{Expected effect of} \\ \text{mutations}}}

>>>>

As in the one-dimensional case, we can go a little further if the noise is sufficiently weak. Let the deterministic trajectory obtained by solving \eqref{nD_det_limit} be given by $\boldsymbol{\alpha}(t) = [\alpha_1(t), \alpha_2(t), \ldots, \alpha_m(t)]^{\mathrm{T}}$.  We can once again track 

@@

&= \sum\limits_{k=1}^{m} y_k\left(\alpha_i\frac{\partial w_i}{\partial x_k}\bigg{|}_{\mathbf{x}=\boldsymbol{\alpha}(t)}\right) + y_iw_i(\boldsymbol{\alpha}) + \mu\sum\limits_{k=1}^{m} y_k\left(\frac{\partial Q_i}{\partial x_k}(\mathbf{x})\bigg{|}_{\mathbf{x}=\boldsymbol{\alpha}(t)}\right)\\
&= y_iw_i(\boldsymbol{\alpha}) + \alpha_iD_{\mathbf{y}}(w_i(\boldsymbol{\alpha})) + \mu D_{\mathbf{y}}(Q_i(\boldsymbol{\alpha}))\label{nD_WNA_directional_derivative_for_replicator_eqns}

@@

\frac{d}{dt}\mathbb{E}[y_i] = \underbrace{w_i(\boldsymbol{\alpha})\mathbb{E}[y_i]}_{\substack{\text{Current fitness of type $i$} \\ \text{at deterministic trajectory $\boldsymbol{\alpha}$} \\ \text{(scaled by expected fluctuation $\mathbb{E}[y_i]$)}}} + \underbrace{\alpha_i\mathbb{E}[D_{\mathbf{y}}(w_i(\boldsymbol{\alpha}))]}_{\substack{\text{Expected change in fitness} \\ \text{ of type $i$ in going from $\boldsymbol{\alpha}$ to $\mathbf{y}$} \\ \text{(scaled by deterministic density $\alpha_i$)}}} + \underbrace{\mu\mathbb{E}[D_{\mathbf{y}}(Q_i(\boldsymbol{\alpha}))]}_{\substack{\text{Expected effect of} \\ \text{mutations}}}
